---
title: "Penguin Analysis"
output:
  html_document: default
  pdf_document: default
date: "2022-12-06"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Purpose of this script: to load penguins dataset, clean it, run a linear regression of body mass against culmen length, create a figure for this test and save it as a series of files for different outputs.



This code does not use a hard file path as its working directory. Instead it will refer to other files within the PenguinProject repository which can be used locally when downloaded together.

#### Define functions used in this code

The following functions are used in the data analysis and visualisation of the dataset. The functions can be found in the 'Functions' folder of the 'PenguinProject' repository.

##### Cleaning

'cleaning' function fixes the column names, removes any empty rows and removes the columns that start with 'delta' and the column called 'comments'.

'remove_NA' function subsets the data to only include penguins that are not NA for both body mass and culmen length.

```{r}
source("functions/cleaning.r")
#includes 'cleaning' and 'remove_NA' functions
```


##### Plotting

'plot_regression' function plots a scatterplot of body mass against culmen length and overlays a linear model.

```{r}
source("functions/plotting.r")
#includes 'plot_regression' function
```


##### Saving

'save_reg_png' function saves the plot as a png when defining the size, resolution, and scaling.

'save_reg_svg' function saves the plot as a svg when defining the size and scaling.

Note that these functions have the 'plot_regression' function embedded into them.

```{r}
source("functions/saving.r")
#includes 'save_reg_png' and 'save_reg_svg' functions
```


#### Load required packages

```{r}
library(palmerpenguins)
library(ggplot2)
suppressPackageStartupMessages(library(janitor))
suppressPackageStartupMessages(library(dplyr))
library(tidyr)
library(ragg)
library(svglite)
```


#### Load the Penguin dataset

```{r}
penguins_raw <- read.csv('data_raw/penguins_raw.csv')
```


#### Clean the data

Use the 'cleaning' function to fix column names, remove empty rows and remove columns that start with 'delta' and 'comments' as we cannot analyse these.

```{r}
penguins_clean <- cleaning(penguins_raw)
```


Save the cleaned data.

```{r}
write.csv(penguins_clean, "data_clean/penguins_clean.csv")
```


#### Linear Regression

We are testing whether body mass is a significant predictor of culmen length in penguins.

Subset the data to include just culmen length and mass without any N/A values using the 'remove_NA' function.

```{r}
penguins_culmen_and_mass <- remove_NA(penguins_clean)
```

Create a linear model with body mass as the explanatory variable and culmen length as the response variable.

```{r}
penguinmod1 <- lm(culmen_length_mm ~ body_mass_g, penguins_culmen_and_mass )
```

Run ANOVA analysis on our linear regression to find if body mass significantly explains culmen length.

```{r}
anova(penguinmod1)
```
As p > 0.05 in this linear model, we can support the hypothesis that body mass significantly explains culmen length in these penguin species.

##### Plot the linear regression with 95% confidence interval.
```{r, message=FALSE}
culmen_mass_plot <- plot_regression(penguins_culmen_and_mass)
culmen_mass_plot
```

Figure 1: A scattergraph of body mass against culmen length. A fitted regression is overlayed on the graph. The grey shaded area is the 95% confidence interval.

#### Save Figure

Save the plot for a report using our 'save_reg_png' function, we specify an appropriate size, resolution and scale for use in a report.
```{r, results='hide', message=FALSE}

save_reg_png(penguins_culmen_and_mass, "figures/fig01_report.png", 
                      size = 15, res = 600, scaling = 1)

```

The same function with new parameters is used for use in a presentation.
```{r, results='hide', message=FALSE}
save_reg_png(penguins_culmen_and_mass, "figures/fig01_presentation.png", 
                      size = 15, res = 600, scaling = 1.4)
```

The same function with new parameters is used for use in a poster.
```{r, results='hide', message=FALSE}
save_reg_png(penguins_culmen_and_mass, "figures/fig01_poster.png", 
                      size = 30, res = 600, scaling = 2.8)
```

Save the plot as a vector using the 'save_reg_svg' function, resolution need not be specified using this function.
```{r, results='hide', message=FALSE}
save_reg_svg(penguins_culmen_and_mass, "figures/fig01_vector.svg", 
                      size = 15, scaling = 1)
```







